---
title: "Section Assignment 1"
author: ""
date: ""
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

This week, your assignment is to familiarize yourself with the Rstudio environment and successfully knit an R markdown file. You will also learn how to save these files as a PDF and upload them to Canvas to be graded. Each week during these section times, you will work in pairs or small groups to complete the tasks. 


First, be sure the author line above (line 3) contains your full name 

  (for example, my line 3 reads: 'author: "Liam Mueller" ') 
  
Second, change the date in line 4 to today's date
  
  (for example, my line 4 reads: 'date: "12/31/2021" ') 


Second, be sure not to change anything other than your name in the author line (line 3) and the date in the date line (line 4) in lines 1-12 as it will change how the file knits and is saved. 

Answer the following question by typing the appropriate code in the code box (denoted by ``` signs).



Question:

1. By ancient tradition (1967 or 1972 or 1978 depending on who you ask), the first line of code a programmer writes when learning to code is an attempt to print the statement: "Hello, World!". If this is your first time programming, please print the statement: "Hello, World!" on line 37. If you are coming into this class with prior programming experience, please print a statement describing your past coding experience on line 37, then see if you can help other students around you with your additional coding experience!

```{r}


```

Once you are done, click the "Knit" button above(It looks like a blue ball of yarn). Save the file with your name and the week number in the file name: 

  (for example: "Liam_Mueller_Section_1").  
  
Then upload the pdf file to canvas/gradescope under the Week 1 Section Assignment before the deadline. 

And that is it for this week!
